Fastest Paths Between Points in $\R^2$ for a Particle Moving with Constant Speed and Bounded Centripetal Acceleration

### 2013-10-10 conclusion.tex
Added more to intro.
Small changes to conclusion.

>>>>

\subsection{Future Work}

Future work should first involve a rigorous proof of . Furthermore, the special case of optimal paths for when the particle is \"too close\" to the final point, should be researched. This special case is very interesting, because it involves moving away from the final point for a period of time, and the shape of the optimal path does not appear to be intuitive.

To make the problem more general, one could extend the problem to include particles moving with varying speed. Equations \ref{eq:r-diff} - \ref{eq:r-diff-2} still apply in this case, but the analysis is more complex in this case.

### 2013-10-10 introduction.tex
Small changes to conclusion.

@@

In physics and computational geometry the problem of finding the shortest path given constraints on the radius of curvature.

There are many instances where finding optimal paths is important. For instance, a race car driver wants to know the optimal way to get from one point to another so that he can win his race. For another example, imagine you are sending a spacecraft to a particular destination in space and would like the optimal means of getting there.

The space example directly motivates our model problem. On a spacecraft, there is a limited amount of acceleration that is possible (provided by thrusters). One would potentially like to navigate to a planet while moving around obstacles.

Understanding how to find optimal trajectories will provide greater insight into solving problems like these.

\subsection{Problem Description}

Given a particle, $p$, moving in $\R^2$, we set constraints on its position, velocity . The particle has a starting position and velocity, and must navigate in $\R^2$ to reach a final position. The goal is to minimize the total time spent navigating to the final position.

>>>>

\subsection{The Problem}

Given a particle that starts at a position in $\R^2$, has initial velocity, and has constrained acceleration, what is the path in $\R^2$ for the particle to take to reach a final position in $\R^2$, such that the time spent traveling from the starting position to the ending position is minimized. The particle can have restrictions on its position, velocity, and acceleration.

@@

The problem of finding a fastest path for a particle with constant speed is usefull because it is identical to the problem of finding a shortest path in $\R^2$ given constraints on the radius of curvature of the path, which appears in fields such as physics and computational geometry. One could imagine a race car driver navigating a racetrack with constant speed, as a physical example of the problem.

This example directly motivates our problem, because a racecar driver has a turning rate limited by friction.

Understanding how to find optimal paths for a particle with constant speed moving in $\R^2$ will provide insight into solving even more complicated problems with varying speed and more dimensions such as spacecraft trajectories.

### 2013-10-10 path_theorems.tex
Small changes to conclusion.

Now that we have defined a coordinate system, we are ready to think about particle paths.

\begin{lemma}
  Given a particle, $p$, with constant speed and bounded centripetal acceleration, $\abs{a_c(t)} \le \bar{a_c}$, the minimum radius of curvature of the particle's trajectory is given by

@@

\begin{proof}
This comes directly from the definition of radius of curvature for a particle without any tangential acceleration:

@@

and the fact that

@@

Knowing this fact, the simplest problem to tackle first is one with infinite centripetal acceleration, which means that a particle can change its direction of motion to an arbitrary direction infinitely fast, $\frac{d\abs{\theta}}{dt}$ is unbounded. We now tackle the problem of finding a fastest path from $\bvec{X_1}$ to $\bvec{X_2}$. Working in our modified polar coordinate system centered on $\bvec{X_2}$, it is clear, that our goal is achieved when $r(t) = 0$. Furthermore, it is important that $\theta(t)$ approach $0$ as $r(t) \to 0$, otherwise, the particle will pass by $\bvec{X_2}$. Equation will grow exponentially fast as $r(t) \to 0$, which can be seen in equation (\ref{eq:r-diff}).

@@

Without loss of generality, we can define a cartesian coordinate system, where the origin is at $\bvec{X_1}$ and the positive x-axis passes through $\bvec{X_2}$. Let us say that the two points in our cartesian coordinate system become $(0,0)$ and $(x_2, 0)$ respectively.

@@

To find the distance $l_x(t)$ travelled in the $\bhat{x}$ direction, we can use the relation:
\begin{align}
  l_x(t) &= \int_0^t v_x(t_2) dt_2 \\
         &= \int_0^t \int_0^{t_1} a_x(t_1) dt_1 dt_2
\end{align}

>>>>

Now that we have defined a coordinate system, we are ready to think about particle paths. Working in our modified polar coordinate system centered on $\bvec{X_2}$, it is clear, that our goal is achieved when $r(t) = 0$. Furthermore, it is important that $\theta(t)$ approach $0$ as $r(t) \to 0$, otherwise, the particle will pass by $\bvec{X_2}$.

\begin{lemma} \label{lem:r-curve}
  Given a particle, $p$, with constant speed and bounded centripetal acceleration, $\abs{a_c(t)} \le \bar{a_c}$, the minimum radius of curvature of the particle's trajectory is given by

@@

This comes directly from the definition of radius of curvature for a particle without any tangential acceleration:

@@

and

@@

With Lemma \ref{lem:r-curve} in mind, the simplest problem to tackle first is the fastest path from $\bvec{X_1}$ to $\bvec{X_2}$ where centripetal acceleration is unbounded, which means that a particle can change its direction of motion to an arbitrary direction infinitely fast, $\frac{d\abs{\theta}}{dt}$ is unbounded. The following Theorem

@@

Without loss of generality, we can define a Cartesian coordinate system, where the origin is at $\bvec{X_1}$ and the positive x-axis passes through $\bvec{X_2}$. Let us say that the two points in our Cartesian coordinate system become $(0,0)$ and $(x_2, 0)$ respectively.

@@

To find the distance $l_x(t)$ traveled in the $\bhat{x}$ direction, we can use the relation:
\begin{align}
  l_x(t) &= \int_0^t v_x(t_2) dt_2 \\
         &= \int_0^t \int_0^{t_1} a_x(t_1) dt_1 dt_2
\end{align}
